\chapter{Methodologies and Validation}\label{chap:method_and_validation}

This chapter details the practical methodologies and numerical validation 
that form the foundation for the investigations in this dissertation. Having 
presented the theoretical groundwork in the preceding chapter, we now 
turn to the numerical implementations.

This chapter is organised into 3 main parts. First, 
we present the MLMC algorithm that is implemented throughout this work.


Second, we focus on the Stochastic Heat Equation (SHE). We begin
by defining the specific SHE problem to be analysed and 
deriving the explicit finite difference scheme 
used for its numerical solution. We then 
detail the MLMC implementation,
including the three distinct noise coupling strategies to be investigated. 
To validate this implementation, we apply the method to compute several 
distinct quantities of interest: first and second moments of the solution's
Fourier modes and its total energy. We first derive 
theoretical convergence values and rates for these quantities, and then demonstrate that
our implementation aligns with these.

Third, we address the Dean-Kawasaki equation. Following 
the approach of Cornalba and Fischer \cite{cornalba2025multilevel}, 
we present the finite difference scheme, MLMC implementation and target 
quantity of interest. 
We finally validate our implementation
by demonstrating its convergence and alignment with the results of Cornalba and Fischer.

By establishing and then rigorously verifying our numerical methods, this chapter 
provides the necessary groundwork for the performance analysis that follows. 

\input{sections/2_methods_mlmc}
\input{sections/2_methods_she}
